---
title: "Homework 3"
subtitle: "Mahitha Jangeti"
output: github_document
---
Mahitha Jangeti
10/13/25
```{r setup}
library(tidyverse)
```

## Problem 1
**Loading dataset and creating plot**

```{r}
library(p8105.datasets)
data("instacart")

instacart_df = instacart
```

```{r}
instacart_df %>% 
  group_by(aisle) %>% 
  summarize(n_obs = n()) %>% 
  mutate(aisle_rank = min_rank(desc(n_obs))) %>% 
  arrange(aisle_rank) %>% 
  filter(n_obs > 10000) %>% 
  ggplot(aes(x = aisle, y = n_obs)) +
  geom_col() +
  coord_flip() +
  labs(
    title = "Number of Items Ordered in each Aisle",
    x = "Aisle Name",
    y = "Number of Items Ordered"
  )
```

**Creating Tables Based on Aisle**

```{r}
instacart_df %>% 
  filter(aisle %in% c("baking ingredients")) %>% 
  group_by(aisle, product_name) %>% 
  summarize(n_baking = n()) %>% 
  mutate(baking_rank = min_rank(desc(n_baking))) %>% 
  arrange(baking_rank) %>% 
  filter(baking_rank <= 3)

instacart_df %>% 
  filter(aisle %in% c("dog food care")) %>% 
  group_by(aisle, product_name) %>% 
  summarize(n_dog = n()) %>% 
  mutate(dog_rank = min_rank(desc(n_dog))) %>% 
  arrange(dog_rank) %>% 
  filter(dog_rank <= 3)
  
instacart_df %>% 
  filter(aisle %in% c("packaged vegetables fruits")) %>% 
  group_by(aisle, product_name) %>% 
  summarize(n_packaged = n()) %>% 
  mutate(packaged_rank = min_rank(desc(n_packaged))) %>% 
  arrange(packaged_rank) %>% 
  filter(packaged_rank <= 3)
```

**Top three products in baking ingredients are light brown sugar, pure baking soda, and cane sugar. Top three products in dog food care are snack sticks chicken & rice recipe dog treats, organix chicken & brown rice recipe, and small dog biscuits. Lastly, the top three products in packaged vegetables fruits are organic baby spinach, organic raspberries, and organic blueberries.** 


**Creating Table for Pink Lady Apples and Coffee Ice Cream** 

```{r}
instacart_df %>% 
  filter(product_name %in% c("Pink Lady Apples", "Coffee Ice Cream")) %>% 
   mutate(
     order_dow = recode(order_dow,
                  `0` = "Monday",
                  `1` = "Tuesday",
                  `2` = "Wednesday",
                  `3` = "Thursday",
                  `4` = "Friday",
                  `5` = "Saturday",
                  `6` = "Sunday")) %>% 
  group_by(product_name, order_dow) %>% 
  rename("Product Name" = product_name) %>% 
  summarize(mean_hour = mean(order_hour_of_day, na.rm = TRUE)) %>% 
  pivot_wider(
    names_from = order_dow,
    values_from = mean_hour
  ) %>% 
  knitr::kable(digits = 1)
```

****There are 134 aisles with most items order from the fresh vegetables, fresh fruits, and packed vegetables fruits aisles. **

## Problem 2

```{r}
zori_df = 
  read_csv("./zillow_data/zori.csv", na = c("NA", ".", "")) %>% 
  janitor::clean_names() %>% 
  relocate(county_name) %>% 
  select(-state, county = county_name) %>% 
  mutate(county = county %>% 
            str_replace("New York County", "New York") %>% 
            str_replace("Kings County", "Kings") %>%
            str_replace("Queens County", "Queens") %>%
            str_replace("Bronx County", "Bronx") %>%
            str_replace("Richmond County", "Richmond")) %>% 
  rename(zip_code = region_name) %>% 
  pivot_longer(
    x2015_01_31:x2024_08_31,
    names_to = "date",
    values_to = "zillow observed rent index")
  knitr::kable(head(zori_df, 8))


neighborhood_df = 
  read_csv("./zillow_data/zip_codes.csv", na = c("NA", ".", "")) %>% 
  janitor::clean_names() %>% 
  relocate(county, neighborhood)
  knitr::kable(head(neighborhood_df, 8))
```

**Merging Datasets**

```{r}
zillow_df = 
  left_join(zori_df, neighborhood_df, by = c("zip_code", "county"))%>%
  janitor::clean_names() %>% 
  relocate(county, neighborhood, zip_code, zillow_observed_rent_index)
  knitr::kable(head(zillow_df, 8))
```

**Solving Problem 2**

```{r}
zip_counts =
  zillow_df %>% 
  group_by(zip_code) %>% 
  summarize(n_obs = n()) %>% 
  filter(n_obs == 116)

zip_counts2 =
  zillow_df %>% 
  group_by(zip_code) %>% 
  summarize(n_obs = n()) %>% 
  filter(n_obs < 10)
```

```{r}

```


** 149 zip codes are observed 116 times and 0 are observed less than 10 times